Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution

### 2018-02-16 main.tex
edited all of appendix but still thinking about last comment

To aid in comparison to other works, we define our ISM phases here, as adopted from \citet{Draine2011} Table 1.3 and used consistently throughout this analysis. By construction, these phases are mutually exclusive. In the below, $f_{\rm H_2}$ is the molecular hydrogen fraction by abundance.

\begin{enumerate}
\item Hot Ionized Medium (HIM): $T \geq 10^{5.5}$ K
\item Warm Ionized Medium (WIM): $10^{4}~\rm{K} \leq T < 10^{5.5}~\rm{K} $
\item Warm Neutral Medium (WNM): $10^{2}~\rm{K} \leq T < 10^{4}~ \rm{K}$
\item Cold Neutral Medium (CNM): $T < 10^2$ K, $f_{\rm H_2} \leq 0.5$
\item Molecular: $T < 10^2$, $f_{\rm H_2} > 0.5$
\end{enumerate}

@@

The ionizing radiation photon energies are taken as the average ionizing photon energy for a black body of the star's given T$_{\rm eff}$ obtained from the PARSEC grid. A more accurate approach to compute this energy would convolve the full stellar spectrum and the frequency-dependent absorption cross section. We tested this approach using the frequency-dependent photoionization cross sections from \citet{1996ApJ...465..487V}.\footnote{Source code containing the analytic fits given in \citet{1996ApJ...465..487V} was obtained from \texttt{http://www.pa.uky.edu/~verner/photo.html}} The black body approximation is accurate to within 5\%, yet substantially easier to compute on the fly as integrals over the black body spectrum can be expressed as infinite series that rapidly converge to high precision. Unlike for the ionizing radiation, we assume constant FUV and Lyman-Werner band energies for each star, at 9.8 eV and 12.8 eV respectively.

@@

The left panel of Figure ~\ref{fig:SN histogram} gives the distributions of the peak and average ISM number densities in the SN injection regions for each SNe in the simulations presented in this work. As shown, a majority explode in regions at substantially lower densities than the star formation threshold of 200~cm$^{-3}$. For most SNe, $n_{\rm max} \le 1.0$~cm$^{-3}$, which is due to the substantial pre-SN feedback included in our simulations. The right panel of Figure ~\ref{fig:SN histogram} gives the fractional distribution of the calculated radius of the pressure-driven snowplow ($R_{\rm PDS}$) phase of the Sedov-Taylor expansion, adopting the definitions from \citet{Simpson2016}
\[
R_{\rm PDS} = 
\begin{cases*}
49.3E^{1/4}_{51}n^{-1/2}_{o} & if  Z $\le$ 0.01 \\
18.5E^{2/7}_{51}n^{-3/7}_{o}Z^{-1/7} & if  Z $\ge$ 0.01 \\
\end{cases*}
\]
where $E_{51}$ is the injection energy in units of 10$^{51}$~erg, $n_{o}$ is the number density of the medium, and $Z$ is the metallicity in units of $Z_{o}$. \citet{Simpson2016} found that $R_{\rm PDS} > 4.5 \Delta x$ is needed to resolve the SN explosion with thermal energy injection alone, assuming uniform density in the injection region. In practice, the injection region is never uniform. We give $R_{\rm PDS}$ assuming uniform density at both the average and maximum number density in the injection region of each SN. As shown, a majority of SNe are resolved (to the right of the 4.5$\Delta x$ line), but that up to 10\% are somewhat under-resolved. In general, these are SN explosions from the most massive (i.e. most prompt) stars. We do not expect that resolving these SNe will dramatically alter our results, but note that simulations presented in future work will examine lower mass galaxies at higher resolution than those presented here. In those, the SNe are more resolved, but we note that using our feedback model at resolutions much less than 2 pc will require adopting a different injection mechanism.

@@

\section{Resolution Study}
\label{appendix:resolution study}
We demonstrate the effect of varying resolution on three different results in the evolution of our galaxy model in Figure~\ref{fig:resolution_study}, comparing the star formation rate (left), mass evolution (middle), and metal ejection/retention fractions (right) for our high resolution simulations (solid) as compared to two runs at 3.6 pc with different initial SN driving rates, and a final run at 7.2 pc. The physics in each of these simulations remains the same. However, our star formation algorithm is tied to resolution, since star formation proceeds in gas at densities where the Jeans mass becomes unresolved at the maximum resolution. A factor of two decrease in resolution translates to a factor of four decrease in the star formation density threshold, from 200~cm$^{-3}$ in our fiducial runs, to 50~cm$^{-3}$ and 12.5~cm$^{-3}$ in our 3.6~pc and 7.2~pc resolution runs respectively.

\begin{figure*}
\centering
\includegraphics[width=0.33\linewidth]{sfr_resolution_study.png}
\includegraphics[width=0.33\linewidth]{mass_evolution_resolution.png}
\includegraphics[width=0.33\linewidth]{metal_retention_resolution.png}
\caption{A comparison of the SFR (left), disk gas mass (middle), and metal retention (right) evolution for our fiducial run (solid), 3.6~pc run (dotted), and 7.2~pc run (dash-dotted). We also include a 3.6~pc run with an initial SN driving rate increased by a factor of 2 (dashed). The line colors in the middle panel are the same as in Figure~\ref{fig:sfr_mass_evolution}, where the total disk mass is shown in black, cold gas mass in blue, H$_2$ mass in orange, and stellar mass in red.}
\label{fig:resolution}
\end{figure*}

As shown in these diagrams, resolution certainly changes the evolution, but not dramatically. We argue that these differences are within the stochastic variation one would expect from simulations with stochastic star formation, but point has not been confirmed. Clearly, the initial SFR over the first 100~Myr is different in each simulation; it is larger in each of the lower resolution runs. This is because the initial SN driving is less resolved, leading to over cooling and an increase in the initial star formation rate. This is much less an issue for the SN from real star particles as the pre-SN feedback (winds and radiation) are capable of decreasing the local density enough to resolve the Sedov radius. To test this, we re-ran the 3.6~pc run with an increase in the initial SN driving rate by a factor of 2, which subsequently lowers the initial SFR.

@@

Finally, each of the lower resolution runs retain a surprisingly similar fractions of their produced metals in their disks (black curves in right panel of Figure~\ref{fig:resolution}), but this is about a factor of 2 below what is found in the fiducial run (5\% vs. 10\%). The amount that escapes the halo is significantly more, however. This is an interesting difference, as the resolution decrease primarily affects the disk of the galaxy; little to none of the halo volume is refined to the highest resolution, and so should be unaffected by this change. \textbf{I'm a bit puzzled by this difference...}

>>>>

To aid in comparison to other works, we define our ISM phases here, as adopted from \citet{Draine2011} Table 1.3 and used consistently throughout this analysis. By construction, these phases are mutually exclusive. In the below, $f_{\rm H_2}$ is the molecular hydrogen fraction
%mm2 by abundance, [to avoid the ambiguity of fraction by proton number or total mass]
   of the total gas mass.

\begin{enumerate}
\item Hot Ionized Medium (HIM): $T \geq 10^{5.5}$ K
\item Warm Ionized Medium (WIM): $10^{4}~\rm{K} \leq T < 10^{5.5}~\rm{K} $
\item Warm Neutral Medium (WNM): $10^{2}~\rm{K} \leq T < 10^{4}~ \rm{K}$
\item Cold Neutral Medium (CNM): $T < 10^2$ K, $f_{\rm H_2} \leq 0.5$
\item Molecular: $T < 10^2$~K, $f_{\rm H_2} > 0.5$
\end{enumerate}

@@

The ionizing radiation photon energies are taken as the average ionizing photon energy for a black body of the star's given T$_{\rm eff}$ obtained from the PARSEC grid. A more accurate approach to compute this energy would convolve the full stellar spectrum and the frequency-dependent absorption cross section. We tested this approach using the frequency-dependent photoionization cross sections from \citet{1996ApJ...465..487V}.\footnote{Source code containing the analytic fits given in \citet{1996ApJ...465..487V} was obtained from \texttt{http://www.pa.uky.edu/~verner/photo.html}} The blackbody approximation is accurate to within 5\%, yet substantially easier to compute on the fly, as integrals over the blackbody spectrum can be expressed as infinite series that rapidly converge to high precision. Unlike for the ionizing radiation, we assume constant FUV and Lyman-Werner band energies for each star, at 9.8 eV and 12.8 eV respectively.

@@

The left panel of Figure ~\ref{fig:SN histogram} gives the distributions of the peak and average ISM number densities in the SN injection regions for each SN in the simulations presented in this work. As shown, a majority explode in regions at substantially lower densities than the star formation threshold of 200~cm$^{-3}$. For most SNe, $n_{\rm max} \le 1.0$~cm$^{-3}$, which is due to the substantial pre-SN feedback included in our simulations. The right panel of Figure ~\ref{fig:SN histogram} gives the fractional distribution of the calculated radius of the pressure-driven snowplow ($R_{\rm PDS}$) phase of the Sedov-Taylor expansion, adopting the definitions from \citet{Simpson2016}
\begin{equation}
R_{\rm PDS} = 
\begin{cases*}
49.3E^{1/4}_{51}n^{-1/2}_{o} & if  Z $\le$ 0.01 \\
18.5E^{2/7}_{51}n^{-3/7}_{o}Z^{-1/7} & if  Z $\ge$ 0.01 \\
\end{cases*}
\end{equation}
where $E_{51}$ is the injection energy in units of 10$^{51}$~erg, $n_{o}$ is the number density of the medium, and $Z$ is the metallicity in units of $Z_{o}$. \citet{Simpson2016} found that $R_{\rm PDS} > 4.5 \Delta x$ is needed to resolve the SN explosion with thermal energy injection alone, assuming uniform density in the injection region. In practice, the injection region is never uniform. We give $R_{\rm PDS}$ assuming uniform density at both the average and maximum number density in the injection region of each SN. As shown, a majority of SNe are resolved (to the right of the 4.5$\Delta x$ line), but up to 10\% are somewhat under-resolved. In general, these are SN explosions from the most massive (i.e. most prompt) stars. We do not expect that resolving these SNe will dramatically alter our results, but note that simulations presented in future work will examine lower mass galaxies at higher resolution than those presented here. In those, the SNe are more resolved, but we note that using our feedback model at resolutions much less than 2 pc will require adopting a different injection mechanism.

@@

\section{Resolution Study}
\label{appendix:resolution_study}

%mm2
    {\bf [This appendix should be cited in the main paper]}
We demonstrate the effect of varying resolution on three different results in the evolution of our galaxy model in Figure~\ref{fig:resolution_study}, comparing the star formation rate, mass evolution, and metal ejection/retention fractions for our high resolution simulations as compared to two runs at 3.6 pc with different initial SN driving rates, and a final run at 7.2 pc. The physics in each of these simulations remains the same. However, our star formation algorithm is tied to resolution, since star formation proceeds in gas at densities where the Jeans mass becomes unresolved at the maximum resolution. A factor of two decrease in resolution translates to a factor of four decrease in the star formation density threshold, from 200~cm$^{-3}$ in our fiducial runs, to 50~cm$^{-3}$ and 12.5~cm$^{-3}$ in our 3.6~pc and 7.2~pc resolution runs respectively.

\begin{figure*}
\centering
\includegraphics[width=0.33\linewidth]{sfr_resolution_study.png}
\includegraphics[width=0.33\linewidth]{mass_evolution_resolution.png}
\includegraphics[width=0.33\linewidth]{metal_retention_resolution.png}
\caption{A comparison of the SFR (left), disk gas mass (middle), and metal retention (right) evolution for our fiducial run (solid), 3.6~pc run (dotted), and 7.2~pc run (dash-dotted). We also include a 3.6~pc run with an initial SN driving rate increased by a factor of 2 (dashed). The line colors in the middle panel are the same as in Figure~\ref{fig:sfr_mass_evolution}, where the total disk mass is shown in black, cold gas mass in blue, H$_2$ mass in orange, and stellar mass in red.}
\label{fig:resolution_study}
\end{figure*}

Figure~\ref{fig:resolution_study} shows that, while decreasing the resolution certainly changes the evolution, it does not do so dramatically. We argue that these differences are within the stochastic variation one would expect from simulations with stochastic star formation.
%mm2, but point has not been confirmed. 
Clearly, the initial SFR over the first 100~Myr is different in each simulation: it is larger in each of the lower resolution runs. This is because the initial SN driving is less resolved, leading to overcooling and an increase in the initial star formation rate. This is much less an issue for the SN from real star particles as the pre-SN feedback (winds and radiation) are capable of decreasing the local density enough to resolve the Sedov radius. To test this, we re-ran the 3.6~pc run with an increase in the initial SN driving rate by a factor of 2, which subsequently lowers the initial SFR.

@@

Finally, each of the lower resolution runs retain a surprisingly similar fractions of their produced metals in their disks (black curves in right panel of Figure~\ref{fig:resolution}), but this is about a factor of 2 below what is found in the fiducial run (5\% vs. 10\%). The amount that escapes the halo is significantly more, however. This is an interesting difference, as the resolution decrease primarily affects the disk of the galaxy; little to none of the halo volume is refined to the highest resolution, and so should be unaffected by this change. \textbf{I'm a bit puzzled by this difference...  }
